---
title: "Minimal Automated Test Assembly Example"
author: "Benjamin Becker"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Minimal Automated Test Assembly Example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`eatATA` efficiently translates test design requirements for Automated Test Assembly (`ATA`) into constraints for a Mixed Integer Linear Programming Model (MILP). A number of efficient and user-friendly functions are available, and the resulting matrix of constrains can be easily transformed to be in line with a MILP solver, like the `GLPK` solver. In the remainder of this vignette we will illustrate the use of `eatATA` using a minimal example.

## Setup
The `eatATA` package can be installed from `CRAN`.  

```{r installation, eval = FALSE}
install.packages("eatATA")
```

As a default solver, we recommend `GLPK`, which is automatically installed alongside this package. If you want to use `Gurobi` as a solver (the most powerful and efficient solver currently implemented in `eatATA`), an external software installation and licensing is required. This means, you need to install the `Gurobi` solver and its corresponding `R` package. A detailed vignette on the installation process can be found [here](https://CRAN.R-project.org/package=prioritizr/vignettes/gurobi_installation.html). 

First, `eatATA` is loaded into your `R` session.

## Example Code
```{r library, message = FALSE}
# loading eatATA
library(eatATA)

items_sim

items_sim[, "IIF"] <- calculateIIF(B = items_sim$difficulty, theta = 0)
nItems <- nrow(items_sim)

itemNumber <- itemsPerFormConstraint(nForms = 1, operator = "=", targetValue = 10, 
                                     itemIDs = items_sim$id)

itemUsage <- itemUsageConstraint(nForms = 1, operator = "<=", targetValue = 1, 
                                 itemIDs = items_sim$id)

testTime <- itemValuesDeviation(nForms = 1, itemValues = items_sim$mean_time,
                    targetValue = 8 * 60, allowedDeviation = 5,
                    relative = FALSE, itemIDs = items_sim$id)


testInfo <- maxConstraint(nForms = 1, itemValues = items_sim$IIF,
                          itemIDs = items_sim$id)


solver_out <- useSolver(list(itemNumber, itemUsage, testTime, testInfo),
                        solver = "GLPK")

inspectSolution(solver_out, items = items_sim, idCol = "id")

```


